---
title: "GSO"
author: "Adam Stuller"
output: pdf_document
---

```{r, include = FALSE}
source('setup.R')
```

# GSO - Ground surface contition

Kategoricka velicina o stave zemi. Zakodovana ako: 

00 = Surface of ground dry (no appreciable amount of dust or loose sand) 

01 = Surface of ground dry (without cracks and no appreciable amount of dust or loose sand and without snow or measurable ice cover) 

02 = Extremely dry with cracks (without snow or measurable ice cover) 

03 = Loose dry dust or sand not covering ground completely (without snow or measurable ice cover) 

04 = Loose dry dust or sand covering more than one-half of ground (but not completely) 

05 = Loose dry dust or sand covering ground completely 

06 = Thin cover of loose dry dust or sand covering ground completely (without snow or measurable ice cover) 

07 = Moderate or thick cover of loose dry dust or sand covering ground completely (without snow or measurable ice cover) 

08 = Surface of ground moist 

09 = Surface of ground moist (without snow or measurable ice cover) 

10 = Surface of ground wet (standing water in small or large pools on surface) 

11 = Surface of ground wet (standing water in small or large pools on surface without snow or measurable ice cover) 

12 = Flooded (without snow or measurable ice cover) 

13 = Surface of ground frozen

14 = Surface of ground frozen (without snow or measurable ice cover) 

15 = Glaze or ice on ground, but no snow or melting snow 

16 = Glaze on ground (without snow or measurable ice cover) 

17 = Ground predominantly covered by ice 

18 = Snow or melting snow (with or without ice) covering less than one-half of the ground 

19 = Snow or melting snow (with or without ice) covering more than one-half of the ground but ground not completely covered 

20 = Snow or melting snow (with or without ice) covering ground completely 

21 = Loose dry snow covering less than one-half of the ground 

22 = Loose dry snow covering at least one half of the ground (but not completely) 

23 = Even layer of loose dry snow covering ground completely 

24 = Uneven layer of loose dry snow covering ground completely 

25 = Compact or wet snow (with or without ice) covering less than one-half of the ground 

26 = Compact or wet snow (with or without ice) covering at least one-half of the ground but ground not completely covered 

27 = Even layer of compact or wet snow covering ground completely 

28 = Uneven layer of compact or wet snow covering ground completely 

29 = Snow covering ground completely; deep drifts 30 = Lose dry dust or sand covering one-half of the ground (but not completely) 

31 = Loose dry snow, dust or sand covering ground completely 

99 = Missing

```{r load, warning = FALSE}
all_data <- read.csv(file= "../data/all.csv")
```

### Centralna poloha dat

Modus - najčastejšia hodnota je 9. 

```{r}
unique(all_data$GSO)

counts <- table(all_data$GSO)
counts / sum(counts)

getmode(na.omit(all_data$GSO)) %>%
  print(cat("Modus: " ))
```


### Barplot 

```{r}
df <- all_data %>%
  dplyr::mutate(
    GSO = map_chr(.$GSO, as.character)
  ) %>% 
  dplyr::filter(!is.na(GSO))

ggplot(df, aes(x = GSO)) + 
  geom_bar(fill = "orange", width = 0.7) +
  xlab("Stav zemi") + 
  ylab("Pocet zaznamenani")
```



